---
title: "Reproducible Research-Project 1"
author: "Sameer Sharma"
date: "March 20, 2018"
output: html_document
---
We first start by reading the CSV file

```{r}

activity_raw<-read.csv("activity.csv",
                       stringsAsFactors=FALSE)
```

Let us ensure that the dates in the dataset are well understood by R as being dates themselves!

```{r}
activity_raw$date <- as.Date(activity_raw$date)
```

What is mean total number of steps taken per day?

Q1:  Calculate total Number of Steps Per Day
```{r}
total_steps_per_day<- tapply(activity_raw$steps,
                             activity_raw$date,sum)
```
Q2:   Histogram of Steps Taken Each Day
```{r}
hist(total_steps_per_day,xlab="Steps Per Day", main = 
       "Histogram of Steps Per Day",col="red")
```
Q3:  Calculate and report the mean and median of the total number of steps taken per day
```{r}
MeanPerDay <- mean(total_steps_per_day, na.rm = TRUE)
MedianPerDay <- median(total_steps_per_day, na.rm = TRUE)
print(MeanPerDay)
print(MedianPerDay)

```
The mean steps per day=10766.19 and 
Median steps per day=10765

What is the Average Daily Pattern?

Q1:  Make a time series plot of the 5-minute interval 
and the average number of steps taken, averaged across all days
```{r}
stepsperint<-tapply(activity_raw$steps,activity_raw$interval,
                    mean,
                    na.rm=T)

plot(as.numeric(names(stepsperint)), 
     stepsperint, 
     xlab = "Interval", 
     ylab = "Steps", 
     main = "Average Daily Activity Pattern", 
     type = "l")
```
Q2:  Which 5-minute interval, on average across all the 
days in the dataset, contains the maximum number of steps?
```{r}
rank<-sort(stepsperint,decreasing = T)
head(rank,1)
```
At 835th interval, max steps is 206.1698

Imputing missing values
Note that there are a number of days/intervals where there are missing values (coded as NA). The presence of missing days may introduce bias into some calculations or summaries of the data.

Q1: What is the number of missing values in the dataset?
```{r}
sum(is.na(activity_raw$steps))
```
There are 2304 missing step values in the raw dataset.

Q2: Devise strategy to fill in missing values
We will use ImputeTS package to fill in missing values
```{r}
require(imputeTS)
activity<-na.interpolation(activity_raw)
```

Q3: Histogram of steps taken each day and mean and median
```{r}

total_steps<-tapply(activity$steps,activity$date,sum)

hist(total_steps, 
     xlab = "Number of Steps", 
     main = 
       "Histogram: Steps per Day 
     (Interpolated for Missing Data)")

MeanPerDay_interpolated <- mean(total_steps, na.rm = TRUE)
MedianPerDay_interpolated <- median(total_steps, na.rm = TRUE)
print(MeanPerDay_interpolated)
print(MedianPerDay_interpolated)
```
Given the interpolation methodology used, we note changes to the mean and median wherein 
                Mean Steps= 9354.23
                Median Steps= 10395

Are there differences in activity patterns between weekdays and weekends?

Q1:  Seperating data into weekdays or weekends
Here we will use ifelse statement
```{r}
activity$day <- ifelse(weekdays
                       (as.Date(activity$date)) == 
                         "Saturday" | 
                         weekdays(as.Date(activity$date)) 
                       == "Sunday", "weekend", "weekday")
```

Q2:  Average Steps/Interval Weekends vs Weekdays We need to also make a  pannel plot
```{r}
stepsperintweekend <- tapply(activity[activity$day 
                          == "weekend" ,]$steps, 
                    activity[activity$day == "weekend" ,]$interval, mean, 
                          na.rm = TRUE)

stepsperintweekday <- tapply(activity[activity$day 
                                      == "weekday" ,]$steps, 
                             activity[activity$day == "weekday" ,]$interval, mean, 
                             na.rm = TRUE)

par(mfrow=c(1,2))

plot(as.numeric(names(stepsperintweekday)), 
     stepsperintweekday, 
     xlab = "Interval", 
     ylab = "Steps", 
     main = "Activity Pattern (Weekdays)", 
     type = "l")


plot(as.numeric(names(stepsperintweekend)), 
     stepsperintweekend, 
     xlab = "Interval", 
     ylab = "Steps", 
     main = "Activity Pattern (Weekends)", 
     type = "l")
```

We can clearly see that weekend activity is higher than weekday activity.






